Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-11-02 manuscript.tex

  \author{Masoud Jasbi\textsuperscript{1}, Brandon Waldon\textsuperscript{1}, \& Judith Degen\textsuperscript{1}}

@@

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Stanford University, Department of Linguistics  }

  \authornote{
    The authors declare that they have no affiliations with, or involvement
    in any organization or entity with any financial interest, or
    non-financial interest in the subject matter or materials discussed in
    this manuscript.

    Correspondence concerning this article should be addressed to Masoud
    Jasbi, Margaret Jacks Hall, Building 460 Rm. 127, Stanford, CA,
    94305-2150. E-mail:
    \href{mailto:masoudj@stanford.edu}{\nolinkurl{masoudj@stanford.edu}}
  }

@@

The past 15 years have seen the rise and development of a bustling and
exciting new field at the intersection of linguistics, psychology, and
philosophy: \emph{experimental pragmatics} (Barner, Brooks, \& Bale,
2011; Bonnefon, Feeney, \& Villejoubert, 2009; Bott \& Chemla, 2016;
Bott \& Noveck, 2004; Breheny, Ferguson, \& Katsos, 2013; Breheny,
Katsos, \& Williams, 2006; Chierchia et al., 2001; De Neys \& Schaeken,
2007; Degen \& Tanenhaus, 2015, 2016; Geurts \& Pouscoulous, 2009;
Grodner, Klein, Carbary, \& Tanenhaus, 2010; Huang \& Snedeker, 2009;
Katsos \& Bishop, 2011; I. A. Noveck \& Reboul, 2008; Noveck \& Posada,
2003; Papafragou \& Tantalou, 2004; Tiel, Miltenburg, Zevakhina, \&
Geurts, 2014; Tomlinson, Bailey, \& Bott, 2013). Experimental pragmatics
is devoted to experimentally testing theories of how language is used in
context. How do listeners draw inferences about the -- often
underspecified -- linguistic signal they receive from speakers? How do
speakers choose between the many utterance alternatives they have at
their disposal?

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled by
  context in a second effortful step if required by context (Bott \&
  Noveck, 2004; Breheny et al., 2006; Degen \& Tanenhaus, 2016; Grodner
  et al., 2010; Huang \& Snedeker, 2009; Politzer-Ahles \& Fiorentino,
  2013; Tomlinson et al., 2013)?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived (Bergen \& Grodner, 2012;
  Bonnefon et al., 2009; Breheny et al., 2013, 2006; Chemla \& Spector,
  2011; De Neys \& Schaeken, 2007; Degen, 2015; Degen \& Goodman, 2014;
  Degen \& Tanenhaus, 2015, 2016; Marneffe \& Tonhauser, 2016; Potts,
  Lassiter, Levy, \& Frank, 2015; Zondervan, 2010)?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  Tiel et al., 2014)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner et al., 2011; Horowitz, Schneider, \& Frank, 2017; Katsos \&
  Bishop, 2011; Musolino, 2004; Noveck, 2001; Papafragou \& Tantalou,
  2004; Stiller, Goodman, \& Frank, 2015)?
\end{enumerate}

@@

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as ``Some, and possibly
all, elephants are mammals'' or \enquote{You got some, and possibly all,
of the gumballs}.

@@

@@

\hypertarget{ref-VanTiel2014}{}
Tiel, B. van, Miltenburg, E. van, Zevakhina, N., \& Geurts, B. (2014).
Scalar diversity. \emph{Journal of Semantics}.
doi:\href{https://doi.org/10.1093/jos/ffu017}{10.1093/jos/ffu017}

@@

>>>>

  \author{Masoud Jasbi\textsuperscript{1}, Brandon Waldon\textsuperscript{2}, \& Judith Degen\textsuperscript{2}}

@@

  \affiliation{
    \vspace{0.5cm}
       \textsuperscript{1} Harvard University, Department of Linguistics\\ 
	\textsuperscript{2} Stanford University, Department of Linguistics }
  \authornote{
    The authors declare that they have no affiliations with, or involvement
    in any organization or entity with any financial interest, or
    non-financial interest in the subject matter or materials discussed in
    this manuscript.

    Correspondence concerning this article should be addressed to Masoud
    Jasbi, Margaret Jacks Hall, Building 460 Rm. 127, Stanford, CA,
    94305-2150. E-mail:
    \href{mailto:masoud_jasbi@fas.harvard.edu}{\nolinkurl{masoudj@stanford.edu}}
  }

@@

The past 15 years have seen the rise and development of a bustling and
exciting new field at the intersection of linguistics, psychology, and
philosophy: \emph{experimental pragmatics} (Barner, Brooks, \& Bale,
2011; Bonnefon, Feeney, \& Villejoubert, 2009; Bott \& Chemla, 2016;
Bott \& Noveck, 2004; Breheny, Ferguson, \& Katsos, 2013; Breheny,
Katsos, \& Williams, 2006; Chierchia et al., 2001; De Neys \& Schaeken,
2007; Degen \& Tanenhaus, 2015, 2016; Geurts \& Pouscoulous, 2009;
Grodner, Klein, Carbary, \& Tanenhaus, 2010; Huang \& Snedeker, 2009;
Katsos \& Bishop, 2011; I. A. Noveck \& Reboul, 2008; Noveck \& Posada,
2003; Papafragou \& Tantalou, 2004; van Tiel, van Miltenburg, Zevakhina, \&
Geurts, 2016; Tomlinson, Bailey, \& Bott, 2013). Experimental pragmatics
is devoted to experimentally testing theories of how language is used in
context. How do listeners draw inferences about the -- often
underspecified -- linguistic signal they receive from speakers? How do
speakers choose between the many utterance alternatives they have at
their disposal?

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled by
  context in a second effortful step if required by context (Bott \&
  Noveck, 2004; Breheny et al., 2006; Degen \& Tanenhaus, 2016; Grodner
  et al., 2010; Huang \& Snedeker, 2009; Politzer-Ahles \& Fiorentino,
  2013; Tomlinson et al., 2013)?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived (Bergen \& Grodner, 2012;
  Bonnefon et al., 2009; Breheny et al., 2013, 2006; Chemla \& Spector,
  2011; De Neys \& Schaeken, 2007; Degen, 2015; Degen \& Goodman, 2014;
  Degen \& Tanenhaus, 2015, 2016; Marneffe \& Tonhauser, 2016; Potts,
  Lassiter, Levy, \& Frank, 2015; Zondervan, 2010)?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  van Tiel et al., 2016)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner et al., 2011; Horowitz, Schneider, \& Frank, 2017; Katsos \&
  Bishop, 2011; Musolino, 2004; Noveck, 2001; Papafragou \& Tantalou,
  2004; Stiller, Goodman, \& Frank, 2015)?
\end{enumerate}

@@

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, e.g. to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as ``Some, and possibly
all, elephants are mammals'' or \enquote{You got some, and possibly all,
of the gumballs}.

\change{Using the rate of \enquote{FALSE} responses in true but underinformative trials as a proxy for implicature calculation is relatively common in experimental pragmatics. For example, in the first study to investigate scalar implicatures experimentally, Noveck (2001) tested adults' and children's interpretations of the scalar items \emph{might} and \emph{some}. The dependent measure in Noveck (2001) was the rate of ``logically correct responses'', i.e. responding ``yes'' to statements such as \emph{Some giraffes have long necks} or \emph{There might be a parrot [in the box]} when there had to be a parrot in the box. He found that children show a higher rate of logically correct responses than adults, and concluded that children interpret scalar items \emph{some} and \emph{might} more logically (i.e. literally). Similarly in another landmark study, Papafragou and Musolino (2003) tested children and adults interpretation of the following set of scalar items: <\emph{two}, \emph{three}>, <\emph{some}, \emph{all}>, and <\emph{finish}, \emph{start}>. The dependent measure in this study was ``the proportion of \emph{No} responses to the puppet's [underinformative] statement''. The study concluded that ``while adults overwhelmingly rejected infelicitous descriptions, children almost never did so''. Furthermore, the study compared the implicature rates across scales and concluded that ``children also differed from from adults in that their rejection rate on the numerical scale was reliably higher than on the two other scales.'' In their final experiment, Papafragou and Musolino (2003) modified their task to invite scalar inferences more easily. They reported that this manipulation resulted in a significantly higher rejection rates. Based on these results, they concluded that children's ability to compute implicatures is affected by the type of scalar item as well as children's awareness of the task's goals. Since these early pioneering studies, many other studies have used the rate of false responses in underinformative trials as their dependent measure including Doran et al (2012), van Tiel et al (2016), Geurts and Pouscoulous (2009), Potts, Lassiter, Levy, and Frank (2016), to name a few.}

@@

\hypertarget{ref-Papafragou2003}{}
Papafragou, A., \& Musolino, J. (2003). Scalar implicature: experiments at the semantics-pragmatics interface. \emph{Cognition}, \emph{86}(3), 253--282.

@@

@@

\hypertarget{ref-VanTiel2014}{}
van Tiel, B., van Miltenburg, E., Zevakhina, N., \& Geurts, B. (2016).
Scalar diversity. \emph{Journal of Semantics}.
doi:\href{https://doi.org/10.1093/jos/ffu017}{10.1093/jos/ffu017}

### 2018-11-02 writeup.Rmd

kable(fixed_effects_table,digits = 2 , caption = "\\label{tab:modeltable}Model parameter estimates and their credible intervals. Rows marked with an asterisk in the evidence column do not contain 0 in the credible interval, thereby providing evidence for an effect.", booktabs = T)
```

>>>>

kable(fixed_effects_table,digits = 2 , caption = "\\label{tab:modeltable}Model parameter estimates and their credible intervals. Rows marked with an asterisk in the evidence column do not contain 0 in the credible interval, thereby providing evidence for an effect.", booktabs = T)

## auxiliary analysis of 3-way interaction

# exhaustive trials
sbq = implicature_rate %>%
  filter(response_type %in% c("binary","quinary") & trial_type == "scalar") %>%
  droplevels()
m.simple <- glmer(implicature ~ definition*response_type - response_type + (1|card) + (1|participant), family="binomial", data=sbq)
summary(m.simple)

# scalar trials
ebq = implicature_rate %>%
  filter(response_type %in% c("binary","quinary") & trial_type == "exhaustive") %>%
  droplevels()
m.simple <- glmer(implicature ~ definition*response_type - response_type + (1|card) + (1|participant), family="binomial", data=ebq)
summary(m.simple)

```
